---
title: "BodyFat Analysis"
author: "Shiwei Cao, Shurong Gu, Yuwei Sun"
date: "February 2, 2018"
output: beamer_presentation
---
```{r setup, include=FALSE}
require(viridis)
highlight = function(x,color='blue')
{
  paste("<b><font color=",color,">",x,"</font></b>",sep="")  
}
blue = function(x) { highlight(x,color='blue')}
green = function(x) { highlight(x,color='forestgreen')}
red = function(x) { highlight(x,color='red')}
```

## Procedure

* **Data preprocessing**: EDA, data cleaning
* **Statistical Modeling**: 
     * Multiple Linear Regression
          * Variable Selection: stepwise method
          * Model Diagnosis
     * Lasso
* **Model selection**:
     * Make predictions on validation set and choose the best model (the one with smallest mse)
* **Model Interpretation**, **"rule of thumb"**
* **Strengths and weaknesses**

## Explore the data
```{r echo=FALSE,fig.align="center"}
library(png)
library(grid)
img <- readPNG("data.png")
 grid.raster(img)
```

## Data Cleaning

```{r, echo = FALSE,fig.width=5, fig.height=4.5}
BodyFat=read.csv("BodyFat.csv",header=T)
suppressMessages(library(tidyverse))
d=data.frame(x=1/BodyFat$DENSITY,y=BodyFat$BODYFAT)
options(repr.plot.width=3, repr.plot.height=3)
ggplot(d,aes(x,y))+geom_point(col="blue")+
   scale_y_continuous(name="Bodyfat")+
   scale_x_continuous(name="Density")+
   theme_bw()+
   theme(
    axis.title.x = element_text(size=14,face="bold"),
    axis.title.y = element_text(size=14,face="bold"),
    axis.text.y = element_text(face="bold",size=12),strip.text = element_text(face="bold",size=12),
    axis.text.x = element_text(face="bold",size=12))
```

##

#####  We decide to remove 3 points for further analysis:
###### The one with BODYFAT=0; the one with HEIGHT=29.5 inches (only 75cm tall); the one with WEIGHT=363.15 pounds
```{r, echo = FALSE,message=FALSE,fig.width=7, fig.height=4.5}
# library(tidyverse);library(plotly)
# options(repr.plot.width=4, repr.plot.height=4)
# p1 <- plot_ly(BodyFat, x = ~BODYFAT) %>% add_histogram()
# p2 <- plot_ly(BodyFat, x = ~WEIGHT) %>% add_histogram()
# p3 <- plot_ly(BodyFat, x = ~HEIGHT) %>% add_histogram()
# p1
# subplot(p2, p3)
p1=ggplot(BodyFat,aes(BODYFAT))+geom_histogram(binwidth = 3,color="white",fill="hotpink")+
   theme_bw()+
   theme(
    axis.title.x = element_text(size=14,face="bold"),
    axis.title.y = element_text(size=14,face="bold"),
    axis.text.y = element_text(face="bold",size=12),strip.text = element_text(face="bold",size=12),
    axis.text.x = element_text(face="bold",size=12))

p2=ggplot(BodyFat,aes(WEIGHT))+geom_histogram(binwidth = 10,color="white")+
   theme_bw()+
   theme(
    axis.title.x = element_text(size=14,face="bold"),
    axis.title.y = element_text(size=14,face="bold"),
    axis.text.y = element_text(face="bold",size=12),strip.text = element_text(face="bold",size=12),
    axis.text.x = element_text(face="bold",size=12))

p3=ggplot(BodyFat,aes(HEIGHT))+geom_histogram(binwidth = 1,color="white")+
   theme_bw()+
   theme(
    axis.title.x = element_text(size=14,face="bold"),
    axis.title.y = element_text(size=14,face="bold"),
    axis.text.y = element_text(face="bold",size=12),strip.text = element_text(face="bold",size=12),
    axis.text.x = element_text(face="bold",size=12))
source("multiplot.R")
multiplot(p1,p2,p3,layout=matrix(data=c(1,2,1,3),nrow=2))
```

## Variable Selection

- Divide the whole data into train (80%) and validation (20%) set
- Train set:
    Use several different methods to choose the best subset of variables
 1.  Mallow's Cp: leaps() in R (do an exhaustive search)
 2.  Stepwise regression based on AIC 
 3.  Stepwise regression based on BIC    
- Validation set: 
    See if the models generalize well on unseen data
    
```{r, echo=FALSE}
mse=data.frame(Model=c("Mallow's Cp","AIC","BIC"),MSE=c(15.069,14.702,13.989))
mse
```

## BIC results
```{r echo=FALSE,fig.align="center"}
library(png)
library(grid)
img <- readPNG("BIC.png")
 grid.raster(img)
```


## Variables selected by BIC:
#### At this point, we come up with a multiple linear regression:


* `r blue ("_Bodyfat_ ~ _Abdomen_ + _Wrist_ + _Weight_")`

```{r echo=FALSE,fig.align="center"}
img <- readPNG("bodyfat.png")
 grid.raster(img)
```

## Prediction on validation set
```{r,echo=FALSE}
BodyFat=BodyFat[-c(39,42,182),]
set.seed(100)
train_index=sample(1:249,size=200,replace=F)
train=BodyFat[train_index,]
test=BodyFat[-train_index,]
model.BIC<-lm(BODYFAT ~ ABDOMEN + WRIST + WEIGHT,data = train)
plot(predict(model.BIC,test[,-c(1,2,3)]),test[,2],pch=23,bg="red",cex=0.8,
    xlab="Predicted Body Fat %", ylab="Real Body Fat %",main="Prediction Plot")
abline(a=0,b=1,col="black",lwd=3)
```


## Model Diagnosis
#### Adjusted R-squared for the final model: 0.7175
#### Linear regression assumptions?
```{r,echo=FALSE}
layout(matrix(c(1,2),nrow=1))
plot(predict(model.BIC),rstandard(model.BIC),pch=23,bg="red",cex=1,
    xlab="Predicted Body Fat %", ylab="Standardized Residuals",main="Standardized Residual Plot")
options(repr.plot.width=7, repr.plot.height=3.5)
abline(a=0,b=0,col="black",lwd=3)
qqnorm(rstandard(model.BIC),pch=23,bg="red",cex=1)
abline(a=0,b=1,col="black",lwd=3)
```

## Still exist multicollinearity?
```{r,echo=FALSE,message=FALSE}
data<-train[,-1]
m2<-lm(BODYFAT ~ WEIGHT + ABDOMEN + WRIST,data = data)
library(car)
```
```{r}
vif(m2)
```

##Another approach: Lasso

$$\min\limits_{\beta} \sum\limits_{i=1}^N (y_i-\beta_0-\beta_1x_i - \dots - \beta_px_p)^2 +\lambda \left|\sum\limits_{j=1}^p \beta_j \right| $$
```{r echo=FALSE,fig.align="center"}
img <- readPNG("lasso.png")
 grid.raster(img)
```


## Conclusion
* Our proposed linear model to predict body fat %:

  (BodyFat %)= $-23.7937+0.8519\times Abdomen-1.2582\times Wrist-0.0735\times Weight$

* Possible rule of thumb:
    + Your *% Bodyfat* = 
    + Your *abdomen* circumference (cm) $\times 0.85$
    + minus *wrist* circumference (cm) $\times 1.26$
    + minus *weight* (lbs) $\times 0.07$
    + minus 24
*  For a normal graduate male student, with circumferences: Abdomen=85cm, Wrist=18cm, Weight=130lbs, his predicted body fat percentage would be around 16.43%. There is a 95% probability that his body fat is between 8.26% and 24.59%.
  
## Strengths and Weaknesses
- Strengths
 1.  Use a separate validation set to avoid overfitting
 2.  Simple, easy to interpret

- Weaknesses
 1.  May lose information using only 200 data points 
 2.  Trade off between simplicity and precision